---
title: "Simple R Functions "
author: ""
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}


a<- c(1:10)
c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
```{r}

funct1b <- function(xVec1,nVec1){
  
  y <- 1:nVec1
  
  f <- 1 + sum(xVec1^y/y)}


hi <- funct1b(3,2)
hi 

```


\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$
```{r}

funct2 <- function(xVec)
  
{ 
 n <- length(xVec)
  for (i in 1:(n-2)){
xVec[i] <- (xVec[i] + xVec[i+1] + xVec[i+2])/3
}
return(xVec[1:(n-2)])
}
f = c(1,2,3,4,5,6)
funct2(f)
```


### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.

```{r}
funct3 <- function(xVec) {
a <- xVec < 0
b <- xVec < 2
c <- xVec >= 2
if (a) {sum(xVec**2 + xVec * 2 + 3)}
if (b) {sum(xVec + 3)}
if (c) {sum(xVec**2 + xVec*4 -7)}
}
funct3(5)

tmpfn1 <- seq(-3, 3, len=10)
plot(tmpfn1, funct3(tmpfn1), type="l")
    
    

```


### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$

```{r}

funct4 <- function(trix){

trix[trix%%2 == 1] <- 2 * trix[trix%%2 == 1]
trix}

  
T <- matrix(
c(2, 4, 3, 1, 5, 7),
nrow=3,
ncol=3,
byrow = TRUE)

T
h <- funct4(T)
h

```


### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$

```{r}
Funct5 <- function(n,k){
gh <- diag(2, nr = 5)
gh[abs(row(gh) - col(gh)) == 1] <- 1
gh 

}


n = c(4:8)
Funct5(n)

```

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.

```{r}

Funct6 <- function(quad)
{
1 + (quad%%360)%/%90
}

d = c(60,40,50,100)
h = Funct6(d)
h

```


### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

```{r}
Funct7 <- function(k,m,y){
  m <- m - 2
  if (m <= 0){
   m <- m +12 
  }
  c <- as.integer(y/100)
  y <-as.integer(y%%100)
  x <- as.integer(2.6*m -0.2)
  f <- (x + k +y+ as.integer(y/4)+as.integer(c/4)-2*c)%%7 +1
  return(f)
  }


h <- Funct7(21,07,1963)
h

```


(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?

```{r}
h <- Funct7(07,21,1963)
h
```
8(a)Suppose $$x_{0} = 1 and x_{1} = 2 and x_{j} = x_{j}-1 + \frac{2}{xj-1}$$
Write a function testLoop which takes the single argument n and returns the first n ??? 1 values of the sequence $${xj}j???0$$ that means the values of x0, x1, x2, . . . , xn???2

```{r}

Funct8 <- function(x){
  
  n <- length(x)
  x <- rep(NA, n-1)
  x[1] <- 1
  x[2] <- 2
  
  for (i in 3:(n-1)) {
    x[i] <- x[i-1] + 2/x[i-1]}
  return(x)
  } 
 
H <- c(1:6)
Funct8(H)

```
b Now write a function testLoop2 which takes a single argument yVec which is a vector. The function should return $$\sum_{j=1}^{n}e^{j}$$ where n is the length of yVec.

```{r}
funct8b <- function(xvec){
  total <- 0;
  for(j in 1:length(xvec)){
    print(total)
    total <- total + xvec[j]**j  
  }
   return(total)
}

H<- (1:10)
funct8b(H)
```


9a Solution of the difference equation $$x_{n} =rx_{n-1}(1-x_{n-1})$$ with starting value x1.
b.Write a function quadmap( start, rho, niter ) which returns the vector $$x_{1},\cdot \cdot \cdot x_{n}$$ where $$x_{k} =rx_{n-1}(1-x_{n-1})$$  and
niter denotes n,
start denotes x1, and
rho denotes r.

Try out the function you have written:
. for r = 2 and 0 < x1 < 1 you should get xn ??? 0.5 as n ??? ???.
. try tmp <- quadmap(start=0.95, rho=2.99, niter=500)
Now switch back to the Commands window and type:
plot(tmp, type="l")
Also try the plot plot(tmp[300:500], type="l")

```{r}
function9a <- function(start,rho,niter){
  xvec <- vector(niter)
  xvec[1] = start
  for(k in 2:n){
    xvec[k] <- rho*xvec[k-1]*(1-xvec[k-1])
  }
  return(xvec)
}

```


(b) Now write a function which determines the number of iterations needed to get $$| xn ??? xn???1 |< 0.02$$.
So this function has only 2 arguments: start and rho. (For start=0.95 and rho=2.99, the answer is
84.)

```{r}

funct9b <- function(strt, rho)
{

eps <- 0.02
x1 <- strt
x2 <- rho*x1*(1 - x1)
n  <- 1
while(abs(x1 - x2) >= eps) {
x1 <- x2
x2 <- rho*x1*(1 - x1)
n  <- n  + 1
}
n
}



funct9b(strt=0.95, rho=2.99)

```



10a Given a vector (x1, . . . , xn), the sample autocorrelation of lag k is defined to be
rk =
$$\frac{\sum_{i=k+1}^{n}\left ( xi-\bar{x} \right )\left ( xi-k-\bar{x} \right )}{\sum_{i=1}^{n}\left ( xi-\bar{x} \right )^{2}}$$
Thus r1 =$$\frac{\sum_{i=k+1}^{n}\left ( xi-\bar{x} \right )\left ( xi-k-\bar{x} \right )}{\sum_{i=1}^{n}\left ( xi-\bar{x} \right )^{2}} = \frac{\left ( x_{2}-\bar{x}\right )\left ( x_{1 -\bar{x}} \right ) +\cdot \cdot \left ( x_{n} -\bar{x} \right )\left ( x_{n}-\bar{x} \right )}{\sum_{i=1}^{n}\left ( x_{i}-\bar{x} \right )^{2}}$$

Write a function tmpFn(xVec) which takes a single argument xVec which is a vector and returns a
list of two values: r1 and r2.
In particular, find r1 and r2 for the vector (2, 5, 8, . . . , 53, 56).

```{r}
funct10a <- function(x){
n <- length(x)
y <- x- mean(x)
div <- sum(y**2)
r1 = 0
r2 = 0

for(i in 2:n){
r1 <- r1 + sum(y[i]*y[i-1])/div
}

for(i in 3:n){
r2 <- r2 + sum(y[i]*y[i-2])/div
}

print(r1)
print(r2)
}

FT <- c(1:4)
funct8b(FT)
```

(b) (Harder.) Generalise the function so that it takes two arguments: the vector xVec and an integer k
which lies between 1 and n ??? 1 where n is the length of xVec.
The function should return a vector of the values (r0 = 1, r1, . . . , rk).
If you used a loop to answer part (b), then you need to be aware that much, much better solutions are
possible-see exercises 4. (Hint: sapply.)

```{r}
funct10b <- function(x,k){
   n <- length(x)
   y <- x- mean(x)
   div <- sum(y**2)
  
   xvec = integer(k)
   for(j in 1:k){
     for(i in (j+1):n){
       xvec[j]  <- xvec[j] + (y[i]*y[i-j])/div
    }
   
   }
  print(xvec)
}

funct10b(c(1,2,3,4),2)


```









